Binomial Rough Analysis

### 2023-10-13 Lahman Analysis/Rough Analysis/BinomialRoughAnalysis.qmd
Added notes
From Meeting with Dr.Parson

```{r}
library(tidyverse)
library(rethinking)
library(Lahman)
```

@@

Batting <- Batting %>% 
  filter(yearID >= 1945) %>% 
  select(playerID, yearID, lgID, AB, AB, HR) %>% 
  mutate(lg = ifelse(lgID=="NL",1,2))
```

@@

```{r}
slim <- list(
  AB = Batting$AB,
  HR = Batting$HR,
  LG = Batting$lg,
  YEAR = Batting$yearID -1944
)

>>>>

```{r}
library(tidyverse)
library(rethinking)
library(Lahman)
library(scales)
```

@@

Batting <- Batting %>% 
  filter(yearID >= 1945) %>% 
  select(playerID, yearID, lgID, AB, AB, HR) %>% 
  mutate(lg = ifelse(lgID=="NL",1,2))

Batting2 <- Batting %>% 
  filter(yearID >=1973 & yearID <=2021)
```

# From Meeting

```{r}
rate <- Batting %>% 
  group_by(yearID) %>% 
  summarize(sHR = sum(HR),
            sAB = sum(AB)) %>% 
  mutate(p = sHR / sAB,
         se = sqrt(p*(1-p)/sAB))
```

```{r}
ggplot(data=rate)+
  aes(x=yearID, y=p)+
  geom_point()+
  scale_y_continuous(labels=label_percent(), limits = c(0,NA))
```

```{r}
#rate2 <- 
Batting2 %>% 
  group_by(lgID) %>% 
  summarize(sHR = sum(HR),
            sAB = sum(AB)) %>% 
  mutate(p = sHR / sAB,
         se = sqrt(p*(1-p)/sAB))
```

```{r}
rate2 <- Batting2 %>% 
  group_by(yearID, lgID) %>% 
  summarize(sHR = sum(HR),
            sAB = sum(AB)) %>% 
  mutate(p = sHR / sAB,
         se = sqrt(p*(1-p)/sAB)) %>% 
  select(yearID, lgID, p) %>% 
  ungroup() %>% 
  pivot_wider(names_from = lgID, values_from = p) %>% 
  mutate(diff = AL - NL)
```

```{r}
ggplot(data=rate2)+
  aes(x=yearID, y=diff)+
  geom_point()
```

```{r}
ggplot(data=rate2)+
  aes(x=diff)+
  geom_histogram()
```

```{r}
Batting3 <-  Batting %>% 
  group_by(yearID, lgID) %>% 
  summarize(sAB = sum(AB), sHR = sum(HR)) %>% 
  mutate(year = yearID-1944,
         lg = case_when(lgID=="NL" ~ 1,
                        lgID=="AL" ~ 2))
```

```{r}
slim2 <- list(
  N = nrow(Batting3),
  z = n_distinct(Batting3$yearID),
  YEAR = Batting3 %>% pull(year),
  AB = Batting3 %>% pull(sAB),
  HR = Batting3 %>% pull(sHR),
  LG = Batting3 %>% pull(lg)
)

str(slim2)
```

# Stan

See Stan file

```{r}
fit <- stan(
  file = "RoughBinomAnalysis.stan",
  data = slim2,
  chains = 4,
  warmup = 1000,
  iter= 5000,
  cores = 4,
  refresh=100
)
```

```{r}
plot(fit)
```

```{r}
print(fit)
```

```{r}
traceplot(fit)
```

```{r}
saveRDS(fit, file="fit.rds")
readRDS("fit.rds")
```

@@

```{r}
slim <- list(
  AB = Batting$AB,
  HR = Batting$HR,
  LG = Batting$lg,
  #Need to subtract by 1944 to make it an index variable instead of a number
  YEAR = Batting$yearID -1944
)
